---
title: 'Investigation #1: Roll Em!'
author: "Archie Phyo & Ella Trinh"
author-title: Authors
format: html
echo: true
embed-resources: true
title-block-banner: true
theme: default
---

```{r}
#| include: false
options(scipen=999)  # turns off scientific notation
set.seed(11032011)  # set your own seed!
```

Suppose you have a standard, fair, 6-sided die. You will roll it again and again until all 6 sides have come up at least once. How long will this take, on average? More generally, if you have a fair $k$-sided die (or cards marked 1 through $k$ that you’re sampling with replacement), what’s the expected number of tries required to see all $k$ numbers?

------------------------------------------------------------------------

Write a simulation function to explore this question. Your function should have at least\* one input: $k$ = the number of sides on your die. Your function should have exactly one output: a single value of the random variable $X$ = number of rolls required to see all $k$ sides.

\*I **strongly** suggest that you read the entire investigation before beginning to write your code! In particular, make your function flexible enough to handle Question 7 by including a second argument.

Write your Build function in the code chunk below. The function will be called `rollem()`.

```{r Build Step}
rollem <- function(k, times = 1){
  counts <- rep(0, k)
  rolls <- 0
  
  while(min(counts) < times){
    outcome <- sample(1:k, size = 1)
    counts[outcome] <- counts[outcome] + 1
    rolls <- rolls + 1
  }
  
  return(rolls)
}
```

Next comes the Repeat step: generate and store at least 100,000 values of the random variable $X$ assuming $k$ = 6.

```{r Repeat Step}
n <- 100000
x <- replicate(n, rollem(k = 6, times = 1))
```

------------------------------------------------------------------------

Use the results of your Build and Repeat steps to answer the following questions. You should only run your code *once* to answer questions 1-3. Don't re-run it for each question!

### 1. Create a well-labeled histogram of the simulated values of $X$. Summarize the distribution (shape, center, spread) in complete sentences.

```{r}
#| fig-height: 4
#| fig-width: 6

hist(x,
     breaks = 30,
     main = "Simulated Distribution of X",
     xlab = "Number of rolls needed to see all 6 sides",
     ylab = "Frequency")
```

<!-- Your written summary goes below this line -->

### 2. Construct and interpret in context a 95% confidence interval for the expected value of $X$ when $k$ = 6.

```{r}
xbar <- mean(x)
sx <- sd(x)
se <- sx / sqrt(n)

ci_lower <- xbar - 1.96 * se
ci_upper <- xbar + 1.96 * se

c(ci_lower, ci_upper)
```

<!-- Your interpretation goes below this line -->






### 3. Construct and interpret in context a 95% prediction interval for a single future value of $X$ when $k$ = 6.

To construct an approximate 95% prediction interval for a single value of a variable, use the 2.5 and 97.5 percentiles of your output vector. The `quantile()` function in R will prove useful. (There exists another formula for prediction intervals, but it requires your distribution to be approximately Normal.)

```{r}

```

<!-- Your interpretation goes below this line -->






------------------------------------------------------------------------

Now you’re going to investigate how the mean of $X$ grows as we increase the number of sides, $k$. Is the mean of $X$ something like $2k$? $3k$? $k^2$?

### 4. Run your code for $k$ = 1, 5, 10, 20, 30, 40, 50, and 60. Use *n* = 1000 for each $k$; otherwise, it'll take a really long time. For each run of 1000, record the sample mean $\bar{x}$. You will need those eight $\bar{x}$ values for Question 5.

Use a *for* loop. You'll need to create a vector of the eight $k$ values. Your *for* loop should output a vector of the eight $\bar{x}$ values.

```{r}



```

### 5. Create a graph with $(\bar{x}\div k)$ on the vertical axis and $k$ on the horizontal axis.

Pay attention: your y-axis values are not the eight $\bar{x}$ values you just calculated, but rather each $\bar{x}$ value divided by the corresponding $k$ value. These quotients are easy to create with the two vectors from the previous step.

If you use the `plot` function in R with the option `type="l"`, you should get a graph with your eight $(k,\bar{x}\div k)$ points connected by line segments. That's what I want. You're also welcome to use more advanced graphics packages, like ggplot2, if you're familiar with them.

```{r}
#| fig-height: 4
#| fig-width: 6


```

### 6. What function does your graph look like to you? Based on that, make an educated guess about an expression for $E(X)$ as a function of $k$.

Pay attention to what your axes represent: the y-axis is *mean divided by k*. Take that into account when making your educated guess.

<!-- No code required. Write your written explanation here. -->






------------------------------------------------------------------------

Let's make the game more interesting: define $Y$ = the number of rolls of a fair 6-sided die required to see all 6 sides at least **twice**. In questions 7 and 8 below, you'll explore the distribution of $Y$.

Use the R code chunk below to simulate at least 100,000 values of $Y$. Ideally, your original simulation function is flexible enough to handle an arbitrary number of times for seeing each side (e.g., once for $X$, twice for $Y$). Otherwise, create a new function here to simulate rolling until you've seen every side twice.

```{r}

```

### 7. Use the output of your code to repeat 1-3 for the random variable $Y$. Don't forget the interpretations!

<!-- You'll have a mix of R code chunks and exposition below -->






### 8. Based on your simulation results, is the mean of $Y$ twice the mean of $X$? Should it be? Why or why not?

<!-- No code required. Write your written explanation here. -->




