---
title: 'Investigation #2: A Real (Computing) Cluster'
author: "Archie Phyo & and Ella Trinh"
format: pdf
echo: true
---

```{r}
#| include: false
options(scipen=999)  # turns off scientific notation
set.seed(305)  # set your own seed!
```

A computing cluster is shared by 250 independent subscribers. Each day, the probability any particular subscriber uses the cluster is 0.3. The number of tasks submitted by a user to the cluster can be modeled by a geometric distribution with a mean of 6.5. The time that it takes the cluster to complete a task can be modeled by an exponential distribution with mean 2.8 minutes. Completion times are also independent.

------------------------------------------------------------------------

Before you write any code, let's reason out some values that you can use later to validate your simulation. These are arithmetic questions; no R code is required. 

### 1. What is the expected number of tasks submitted in a day?

Expected users: \(250(0.3) = 75\)

Each active user submits an average of 6.5 tasks, so

\[
E(X) = 75(6.5) = 487.5
\]

The expected number of tasks submitted in a day is 487.5.

### 2. What is the expected amount of time required to complete all the day's tasks?

Each task takes an average of 2.8 minutes, so

\[
E(Y) = 487.5(2.8) = 1365
\]

The expected amount of time required is 1365 minutes, or 22.75 hours.

------------------------------------------------------------------------

Now you will write code to simulate one day’s worth of tasks, in order to answer the questions below. Notice that the model has three random variables, and they're *nested*: the number of tasks depends on the number of users that day, and the time to complete all tasks depends on the total number of tasks submitted.

Your main function, `cluster()`, will have no inputs; you may hard-code the parameters 250, 0.3, 6.5, and 2.8. Your function must return the values of **two** variables: *X = the total number of tasks submitted*, and *Y = the total amount of time required to complete all tasks*.

After that, the output of your Repeat step should consist of two sets of numbers: the simulated values for *X* and the simulated values for *Y*.

Don't worry about unfinished tasks “bleeding over” into the next day; you are simulating independent days. Equivalently, assume each day starts fresh, with zero tasks queued up.

Two very important notes:

-   Everything in `cluster()` must be **vectorized**. There should be no *for* loops in your code.

-   R's parameterization of the geometric distribution is different. You will need to adjust for it accordingly. You have been warned.

------------------------------------------------------------------------

<!-- Place your Build function and Repeat code in the code chunk below. -->

```{r Build and Repeat Steps}

# Build

cluster <- function(){
  n_users <- rbinom(1, 250, 0.3)
  n_tasks <- rgeom(n_users, 1/6.5) + 1
  total_tasks <- sum(n_tasks)
  total_time <- sum(rexp(total_tasks, 1/2.8))
  return(c(X = total_tasks, Y = total_time))
}


# Repeat

sims <- 100000
sim_data <- replicate(sims, cluster())
tasks <- sim_data["X", ]
times <- sim_data["Y", ]

```

------------------------------------------------------------------------

Use the output of the previous code chunk to answer the following questions. You should only run that code chunk once.

### 3. Create a well-labeled histogram of 100,000 values of the variable *X = number of tasks submitted in a day*. Summarize the distribution (shape, center, spread) in complete sentences.

```{r Histogram of X}
hist(tasks, 
     main = "Daily Tasks Submitted", 
     xlab = "Tasks", 
     col = "lightgray", 
     breaks = 50)
```
<!-- Your written summary goes below this line. -->
The distribution of daily tasks is roughly normal, centered at a mean of `r round(mean(tasks), 2)` tasks with a standard deviation of `r round(sd(tasks), 2)` tasks, indicating substantial day-to-day variation. The middle 50 percent of days range from `r round(quantile(tasks, 0.25), 2)` to `r round(quantile(tasks, 0.75), 2)` tasks.


### 4. Calculate and interpret a 95% confidence interval for the probability that more than 500 tasks are submitted in a day.

```{r CI Prob 1}
prop.test(sum(tasks > 500), length(tasks))$conf.int
```
<!-- Your written interpretation goes below this line. -->
A 95 percent confidence interval for the probability of more than 500 tasks submitted daily is (`r round(prop.test(sum(tasks > 500), length(tasks))$conf.int[1], 4)` to `r round(prop.test(sum(tasks > 500), length(tasks))$conf.int[2], 4)`). This means we are 95 percent confident the true probability falls between these values.


### 5. Calculate and interpret a 95% confidence interval for the mean number of tasks submitted in a day. Also comment on whether this interval is consistent with the common-sense answer from earlier.

```{r CI mu_X}
t.test(tasks)$conf.int
```

<!-- Your written summary goes below this line. -->
A 95 percent confidence interval for the mean daily tasks is (`r round(t.test(tasks)$conf.int[1], 2)` to `r round(t.test(tasks)$conf.int[2], 2)`) tasks. This matches the expected value of 487.5 tasks since that value lies within the interval.


### 6. Calculate a 95% confidence interval for the standard deviation of the number of tasks submitted in a day.

When $n$ is very large and the distribution of $X$ is approximately normal, a 95% CI for the variance $\sigma^2_X$ has endpoints

$$\frac{s^2}{1 \pm 1.96 \sqrt{2/(n-1)}}$$ where $s$ is the sample standard deviation. A 95% CI for $\sigma_X$ results from taking the square root of the two endpoints.

You must **write a function** that takes in a numerical vector and outputs the two endpoints for the 95% CI for $\sigma_X$. Execute your function to print out the CI (inside the code chunk is fine). No interpretation required.

```{r CI sigma_X}
ci_sd <- function(data) {
  n <- length(data)
  s <- sd(data)
  f <- 1.96 * sqrt(2/(n-1))
  sqrt(s^2 / c(1+f, 1-f))
}

ci_sd(tasks)
```

<!-- Your written summary goes below this line. -->
The 95 percent confidence interval for the standard deviation of daily tasks is between `r round(ci_sd(tasks)[1], 2)` and `r round(ci_sd(tasks)[2], 2)` tasks.


### 7. Calculate and interpret a 95% prediction interval for a single future value of the number of tasks submitted in a day.

When the distribution of $X$ is approximately normal, a 95% PI is given by $$\bar{x} \pm 1.96 \cdot s \cdot \sqrt{1 + \frac{1}{n}}$$ This formula is superior to (more reliable than) the quantile method used in the previous investigation, but it's only valid for normal distributions.

You must **write a function** that takes in a numerical vector and outputs the two endpoints for this 95% PI. Make sure you include an interpretation!

```{r PI X}
pi_95 <- function(data) {
  m <- mean(data); s <- sd(data); n <- length(data)
  margin <- 1.96 * s * sqrt(1 + 1/n)
  m + c(-margin, margin)
}

pi_95(tasks)
```

<!-- Your written summary goes below this line. -->
The 95 percent prediction interval for the number of tasks submitted on a future day ranges from `r round(pi_95(tasks)[1], 1)` to `r round(pi_95(tasks)[2], 1)` tasks. We expect that 95 percent of future days will fall within this task count range.



------------------------------------------------------------------------

### 8. Create a well-labeled histogram of 100,000 values of the variable *Y = time to complete a day’s worth of tasks*. Summarize the distribution (shape, center, spread) in complete sentences.

```{r Histogram of Y}
hist(times, 
     main = "Daily Completion Time", 
     xlab = "Minutes", 
"complete the rest"
```

<!-- Your written summary goes below this line. -->





### 9. Calculate and interpret a 95% confidence interval for the probability that a day’s worth of tasks can be completed within 24 hours.

```{r CI Prob 2}

```

<!-- Your written summary goes below this line. -->





### 10. Calculate and interpret a 95% confidence interval for the mean completion time for a day’s worth of tasks. Also comment on whether this interval is consistent with the common-sense answer from earlier.

```{r CI mu_Y}

```

<!-- Your written summary goes below this line. -->





### 11. Calculate a 95% confidence interval for the standard deviation of the time to complete a day’s worth of tasks.

Use your earlier function to perform the calculation. No interpretation required.

```{r}

```

### 12. Calculate and interpret a 95% prediction interval for a single future value of the time to complete a day’s worth of tasks.

Use your earlier function to perform the calculation. Don't forget the interpretation!

```{r}

```

<!-- Your written summary goes below this line. -->





------------------------------------------------------------------------

Finally, since you've simulated *two* variables, it makes sense to explore the *relationship* between them.

### 13. Create a well-labeled *scatterplot* of the two variables. Describe what you see (direction, form, strength) in complete sentences, and explain why this makes sense.

```{r Scatterplot}

```

<!-- Your written summary goes below this line. -->




### 14. Use your simulation to calculate a 95% confidence interval for the probability that more than 500 tasks are submitted in a day *and* that those tasks can be completed within 24 hours.

Note: that's a single event, not two separate questions. No interpretation required.

```{r CI Prob 3}

```
